---
title: "Quality Control"
author:
- name: David Rach
  affiliation: University of Maryland, Baltimore
  email: drach@som.umaryland.edu
package: Luciernaga
output:
  BiocStyle::html_document
abstract: |
  How to set up Luciernaga
vignette: |
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette addresses some of the instrumental quality control functions implemented by Luciernaga. These primarily are focused on Cytek (TM) Aurora and Northern Light cytometers that utilize the SpectroFlo software for acquistion and unmixing. This is due to they were the only spectral cytometers at our institution during the time the author was developing this package. If you would like to contribute your own function to expand repertoire of instruments and manufacturers covered, please send a pull request via GitHub and we can solve this Aurora heavy bias :D

Cytek Aurora instruments use fluorescent beads as part of their daily quality control checks. These result in the production of a QC report visible to the user, with the underlying data stored under the Levy-Jennings tracking tab. When downloaded, this returns a .csv file. However, it is not in a "tidy" data format, which limits it's direct usefulness for exploratory data analysis via functional programming in R. We provide several functions to convert this output into a "tidy" format, and then to visualize the data contained. 

## CytekQCFilePrep

```{r}


```

## CytekQCPlots

```{r}


```






