---
title: "Quality Control"
author:
- name: David Rach
  affiliation: University of Maryland, Baltimore
  email: drach@som.umaryland.edu
package: Luciernaga
output:
  BiocStyle::html_document
abstract: |
  How to set up Luciernaga
vignette: |
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette addresses some of the instrumental quality control functions implemented by Luciernaga. Due to the only Spectral Cytometers at the authors institution being are Cytek (TM) Aurora and Northern Light cytometers that utilize the SpectroFlo software for acquistion and unmixing, there is a heavy bias towards them in terms of the functions presented in this vignette. If you would like to address this by contributing your own function that expands the repertoire of manufacturers and instruments covered by this vignette, please reach out via GitHub with a pull request. 


## CytekQCFilePrep

Cytek Aurora instruments use fluorescent beads as part of their daily quality control checks. These result in the production of a QC report visible to the user, with the underlying data stored under the Levy-Jennings tracking tab. When downloaded, this returns a .csv file. However, it is not in a "tidy" data format, which limits it's direct usefulness for exploratory data analysis via functional programming in R. We provide several functions to convert this output into a "tidy" format, and then to visualize the data contained. 

The first step is to indicate where the saved .csv file containing the Levy-Jennings tracking data from the cytometer is stored. On a local computer, the code to do this would resemble the following:
```{r}
FileLocation <- file.path("C:", "Users", "JohnDoe", "Desktop", "LevyJenningsTracking.csv")
```

For this vignette, we will be accessing a .csv folder that can be found in Luciernaga's extdata folder containing three months of Levy-Jennings Tracking data for a 5 Laser Cytek Aurora (UV16-V16-B14-YG10-R8) for our example. 

```{r Access Package Data}
File_Location <- system.file("extdata", package = "Luciernaga")
CSV_Pattern <- ".csv$"
CSV_Files <- list.files(path=File_Location, pattern=CSV_Pattern)
CSV_Files
```

Since this .CSV file is not is not in a tidy format, `read.csv()` will have issues when attempting to import directly. Instead, `CytekQCPlots()` will attempt to read in each row individually as text and from there pre-process it. 

```{r}
library(Luciernaga)
TidyData <- CytekQCFilePrep(CSV_Files, TrackChange = FALSE)
```

## CytekQCPlots

```{r}


```


## QC Retrieval

```{r}


```







