---
title: "03 Quality Control"
date: "`r BiocStyle::doc_date()`"
author:
- name: David Rach
  email: drach@som.umaryland.edu
  affiliation: University of Maryland, Baltimore
package: "`r BiocStyle::pkg_ver('Luciernaga')`"
output:
  BiocStyle::html_document
bibliography: "`r file.path(system.file('extdata', package='Luciernaga'), 'refs.bib')`"
vignette: |
  %\VignetteIndexEntry{03 Quality Control}
  %\VignettePackage{Luciernaga}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Instrument QC

This vignette covers the `Luciernaga` functions that are involved in assessing instrumental quality control. Currently, this section is tailored
for Cytek Aurora and Northern Light (TM) spectral cytometers that utilize SpectroFlo as their acquisition and unmixing software. The reason for this
platform bias is that these were the only spectral cytometers at the authors home institution when designing the package. If you would like to help us remedy this historical bias and expand the functions to other instruments, feel free to reach out via GitHub! :) 

The following functions can work from either the Levy-Jennings Tracking Reports that are generated following automated instrument QC (exported as .csv
files) or by retrieving the equivalent data stored as keywords within an .fcs files description for a collection of .fcs files across the timespan of interest. 

```{r, echo=FALSE, results = "hide", warning=FALSE}
suppressPackageStartupMessages({
library(Luciernaga)
library(flowCore)
library(flowWorkspace)
library(openCyto)
library(ggcyto)  
library(data.table)
library(dplyr)
library(purrr) 
library(stringr)
library(ggplot2)
library(gt)
library(plotly)
library(htmltools)
})
```

```{r}
library(Luciernaga)
library(flowCore)
library(flowWorkspace)

library(openCyto)
library(ggcyto)  
library(data.table)
library(dplyr)
library(purrr) 
library(stringr)
library(ggplot2)
library(gt)
library(plotly)
library(htmltools)
```

Cytek Aurora instruments use fluorescent beads as part of their daily quality control checks. These beads have complex signatures, with most beads having peak detectors in either the Ultraviolet or Yellow-Green Laser as can be seen below. 

```{r}
File_Location <- system.file("extdata", package = "Luciernaga")
fcs_files <- list.files(File_Location, pattern=".fcs", full.names = TRUE)
QCBeadFile <- fcs_files[grep("After", fcs_files)][[1]]

QCBead_CS <- load_cytoset_from_fcs(QCBeadFile, transformation = FALSE, truncate_max_range = FALSE)
QCBead_GS <- GatingSet(QCBead_CS)

Data <- Luciernaga_QC(x=QCBead_GS[1], subsets="root", sample.name="GUID", SignatureReturnNow = TRUE, Verbose=TRUE,
              unmixingcontroltype="beads", Unstained=TRUE, removestrings=".fcs", stats="median")

Data <- Data %>% mutate(Sample="QCBeads") %>% relocate(Sample, .before=1)

QCBeads <- QC_WhatsThis(x="QCBeads", data=Data, NumberHits=0, returnPlots=TRUE)

QCBeads[[2]]
```
Using the QC beads, the gains for instrument detectors and power to individual lasers are adjusted to maintain a stable MFI across time. These adjustments are available to individual users as part of a QC report, with the underlying data available under the Levy-Jennings tracking tab, with the option to download as a .csv file. Unfortunately, it is not in  a "tidy" data format, limiting our ability for easy exploratory data analysis. We provide a couple functions to convert the .csv output into a "tidy" format and visualize the data contained within. 

The first step is to indicate where the saved .csv file containing the Levy-Jennings tracking data from the cytometer is stored. On a local computer, the code to do this would look similar to the following:
```{r}
FileLocation <- file.path("C:", "Users", "JohnDoe", "Desktop",
                          "LevyJenningsTracking.csv")
```

For this vignette, we will use one of these Levy-Jennings Tracking .csvs from a 5 Laser Cytek Aurora (UV16-V16-B14-YG10-R8), that can be found within Luciernaga's extdata folder. 

```{r}
File_Location <- system.file("extdata", package = "Luciernaga")
CSV_Pattern <- ".CSV$"
CSV_Files <- list.files(path=File_Location, pattern=CSV_Pattern,
                        full.names=TRUE)
CSV_Files
```

## QC_FilePrep

Since this .CSV file is not in a tidy format, `read.csv()` or similar functions will struggle to import correctly. Our function, `QC_FilePrep()` reads in each row individually as text, and then follows up on the pre-processing into a "tidy" format. 

```{r, message=FALSE, warning=FALSE}
TidyData <- QC_FilePrep(CSV_Files, TrackChange = FALSE)
gt(head(TidyData, 5))
```

Individual rows are the individual QC reports, with the columns being the different parameters. Additionally, if the individual parameter exceeded the variance allowed resulting in a QC failure, it is recorded under the corresponding Flag- column. With the data now "tidy", it can be more readily used in R to plot individual detector(s) that might be of interest. 

Additionally, setting `QC_FilePrep()` TrackChange to TRUE will generate a similar output to that of the QC report showing Change from the previous QC report which can provide additional information for instrument monitoring. 

```{r, message=FALSE, warning=FALSE}
TidyData_Track <- QC_FilePrep(CSV_Files, TrackChange = TRUE)
gt(head(TidyData_Track, 3))
```

## QC_Plots

Once we have our data, we can visualize it. This can be done with `QC_Plots()` for individual parameters of interest, or iterarting over the different columns to generate all the plots. Since there are a large number of parameters contained as individual columns, the argument "MeasurementType" will filter the columns for matching character values in the column names. These will be selected from the rest and their data used for the downstream plotting. 

```{r}
head(colnames(TidyData), 20)
```
```{r}
StorageLocation <- file.path("C:", "Users", "JohnDoe", "Desktop")

AvailableBroadMeasurementTypes <- c("Gain", "rCV", "Laser Delay",
                                    "Area Scaling Factor")
SinglePlot <- "UV7-Gain"

TheSinglePlot <- QC_Plots(x = TidyData, MeasurementType = SinglePlot,
                          FailedFlag = TRUE, returntype="patchwork",
                          path=StorageLocation, filename="CytekAurora5L_QC")

TheSinglePlot
```

```{r}
TodaysExample <- c("Gain")

Plots <- QC_Plots(x = TidyData, MeasurementType = TodaysExample,
                  FailedFlag = TRUE, returntype="patchwork",
                  path=StorageLocation, filename="CytekAurora5L_QC")
Plots[1]
```

Setting FailedFlag = TRUE will add a red box on the graph to showcase the date when a particular detector failed the automated QC check by exceeding the allowed variance set for the instrument. 

As with most functions in Luciernaga, `CytekQCPlots()` uses `Utility_Patchwork()` for the .pdf generation. The layout arrangement can be edited by providing additional arguments to generate a desired layout for desired page size. 


## QC Retrieval

In the absence of QC reports from an instrument, we can use .fcs files acquired on the instrument over a desired period of time to retrieve the same information. This is thanks to a lot of this data being stored as description parameters within an .fcs file.

To do this, we will use .fcs files consisting of 3000 events acquired before the DailyQC as part of monitoring. We will first identify the files of interest, and load them into a GatingSet.

```{r}
File_Location <- system.file("extdata", package = "Luciernaga")
FCS_Pattern <- ".fcs$"
FCS_Files <- list.files(path = File_Location, pattern = FCS_Pattern,
                        full.names = TRUE, recursive = FALSE)
QCBeads <- FCS_Files[grep("Before", FCS_Files)]
head(QCBeads)
```
```{r}
MyCytoSet <- load_cytoset_from_fcs(QCBeads, truncate_max_range = FALSE,
                                   transform = FALSE)
MyGatingSet <- GatingSet(MyCytoSet)
MyGatingSet
```

Having the files now loaded into a GatingSet, we can retrieve the information using `QC_Retrieval` function. 

```{r}
SingleSpecimen <- QC_Retrieval(x=MyGatingSet[[1]], sample.name="TUBENAME")

gt(SingleSpecimen)
```

```{r}
AllSpecimens <- map(.x=MyGatingSet, .f=QC_Retrieval, sample.name="TUBENAME") %>%
  bind_rows()

gt(AllSpecimens)
```

Once we have generated these data outputs, we can now send them to `QC_Plots` for visualization of the instrument gain, laser delay and area scaling factors similar to as if we had retrieved the data from the Levy-Jennings Tracking .csv.

```{r}
YellowExample <- c("YG")

Plots <- QC_Plots(x = AllSpecimens, MeasurementType = YellowExample,
                  FailedFlag = FALSE, returntype="patchwork",
                  path=StorageLocation, filename="CytekAurora5L_QC",
                  therows=3, thecolumns=1, width=7, height=9)

Plots[[1]]
```


## QC_GainMonitoring

While `QC_FilePrep()` and `QC_Retrieval()` provide two ways of retrieving information about the individual detectors gains, they only tell part of the instrument story. Gains and Lasers are dynamically changed to maintain a stable MFI over time. One way that our cytometry core has set out to monitor this is acquiring 3000 QC beads immediately before and after running the daily QC, then extracting and visualizing the MFI values in combination with the gain parameters. These before/after .fcs files are processed by the `QC_GainMonitoring()` function. 

```{r}
QCBeads <- FCS_Files[grep("Before|After", FCS_Files)]
head(QCBeads)

BeforeAfter_CS <- load_cytoset_from_fcs(files=QCBeads, transform=FALSE, truncate_max_range = FALSE)

BeforeAfter <- map(.x=BeforeAfter_CS, .f=QC_GainMonitoring, sample.name = "TUBENAME", stats="median") %>% bind_rows()

gt(head(BeforeAfter, 4))
```

The setup for these plots requires a couple additional arguments. The first Metadata designates the column name under which the comparison categories are stored (in this case, Timepoint column, with Before and After designations). The second is we change plotType from the default "individual" to the "comparison" category. 

```{r}
MFIExample <- c("R1")

Plots <- QC_Plots(x = BeforeAfter, MeasurementType = MFIExample,
                  Metadata = "Timepoint", plotType = "comparison",
                  FailedFlag = FALSE, returntype="patchwork",
                  path=StorageLocation, filename="CytekAurora5L_QC")
Plots[[1]]
```
As we can see from the example, looking at the MFI we can see that after daily QC, the recorded MFI across bead samples is being recorded is similar across time, thanks in part to the changes in the individual gain depicted above. 
 
 
# Library Reference Controls

For Cytek instruments that use SpectroFlo, you are able to store unmixing controls in the library for re-use across future experiments. These can be exported out as .XML files, and re-imported to other instruments. Unfortunately, after original acquisition, there is no good way to visualize what the signature of the stored library control is. Whether the library reference control is accurate, contaminated with autofluorescence, or somewhere in between, we are left to parse out from effect on full-stained samples after unmixing. The following functions are designed to extract the signature information from the .XML files to allow for visualization of the underlying normalized signatures. These additionally can be plotted vs. reference signatures to compare how closely they match. 

For an individual user, we would first load the required libraries, and then provide the location where the exported .XML files are stored. An example is provided below. 

```{r, echo=FALSE, results = "hide", warning=FALSE}
suppressPackageStartupMessages({
library(Luciernaga)
library(xml2)
library(tidyr)
library(purrr)
library(dplyr)
library(ggplot2)
})
```

```{r}
library(Luciernaga)
library(xml2)
library(tidyr)
library(purrr)
library(dplyr)
library(ggplot2)
```

```{r}
FolderLocation <- file.path("C:", "Users", "JohnDoe", "Desktop",
                            "LibraryControls5L")
StorageLocation <- file.path(FolderLocation, "Visualized")
```

In our example today, we will be accessing the .XML files stored within the `Luciernaga` packages extdata folder. 
```{r}
Folder_Location <- system.file("extdata", package = "Luciernaga")
XML_Pattern <- ".XML$"
XML_Files <- list.files(path = Folder_Location, pattern = XML_Pattern,
                        full.names = TRUE, recursive = FALSE)
XML_Files
```
## QC_LibraryParse

`QC_LibraryParse()` is the function that will convert the .XML files into "tidy" data, as either plots or a dataframe. This can either be done individually, or in combination with `map()` for all the files. 

By setting the returntype to "plots", we can return each file as a ggplot object. In combination with 
`Utility_Patchwork()` we can generate a .pdf file or an assembled patchwork plot to our desired dimensions. 


```{r}
SinglePlot <- QC_LibraryParse(XML_Files[2], returntype="plots", references=FALSE)

ThePlots <- map(.x=XML_Files, .f=QC_LibraryParse, returntype="plots",
                references=FALSE)

AssembledPlot <- Utility_Patchwork(x=ThePlots, filename="LibraryControls5L",
                  outfolder=NULL, returntype="patchwork", thecolumns=3,
                  therows = 3, width = 7, height = 9)

AssembledPlot[1]
```

The Library Reference Controls featured above where acquired during a training workshop. As you may notice, some library reference controls appear normal, while others have some quality control issues. Profiling the reference controls acquired on our institutions instruments, we noticed that the location the detector gate is placed on the brightness detector plot ties in to whether the y-axis and peak max-value is at 1. When this is misplaced, the stored reference control has a peak greater than 1. 

By setting the argument "references=TRUE", we can append the reference signatures in red to our existing library controls.

```{r}
ThePlots <- map(.x=XML_Files, .f=QC_LibraryParse, returntype="plots",
                references=TRUE)

AssembledPlot <- Utility_Patchwork(x=ThePlots, filename="LibraryControls5L",
                  outfolder=NULL, returntype="patchwork", thecolumns=3,
                  therows = 3, width = 7, height = 9)

AssembledPlot[1]
```

We have attempted to provide coverage for most fluorophores on all Cytek Aurora and Northern Light Instruments. Some fluorophores not present in the internal .csv, a mismatch in the instrument specific fluorophore name can lead to the absence of a reference signature being appended to the individual plot. You can investigate which of these is the case using the `QC_ReferenceLibrary()` function. 

```{r}
QC_ReferenceLibrary(FluorNameContains = "FITC", NumberDetectors=64)
```

Alternatively for `QC_LibraryParse`, we can set the returntype to "dataframe" and return in a "tidy" format. 

```{r}
LibraryData_Single <- QC_LibraryParse(XML_Files[2], returntype="dataframe", references=FALSE)

LibraryData <- map(.x=XML_Files, .f=QC_LibraryParse, returntype="dataframe", references=FALSE) %>% bind_rows()
gt(head(LibraryData, 4))
```

The data from the extracted signatures can then be saved as a .csv file for future reference. 

```{r}
StorageLocation <- file.path("C:", "Users", "JohnDoe", "Desktop",
                             "LibraryControls5L")
FileName <- "ReferenceData.csv"
StorageName <- file.path(StorageLocation, FileName)
# write.csv(TheData, file=StorageName, row.names=FALSE)
```

Please be aware, if the .XML library controls imported originated from different instruments (consequently different laser and detector configurations), 
the `map()` followed with `bind_rows()` can occasionally fail to produce the desired output. A workaround is to take the `map()` outputted rows, sort them based on number of columns, and then bind them into their own data.frames on the basis of shared instruments laser and detector configurations. An example of how we did this is provided below and can be adapted for your individual situation. 

```{r}
#Setting up example
LibraryData <- map(.x=XML_Files, .f=QC_LibraryParse, returntype="dataframe", references=FALSE)

LibraryData[[3]] <- LibraryData[[3]][, -ncol(LibraryData[[3]])]

ncol(LibraryData[[3]]) == ncol(LibraryData[[1]])
```

```{R}
FirstItem <- ncol(LibraryData[[1]])

MisbehavingRows <- which(sapply(LibraryData, function(x) ncol(x) != FirstItem))

if(length(MisbehavingRows) != 0) {
  TheData_filtered <- LibraryData[-MisbehavingRows]
} else {TheData_filtered <- LibraryData}

TheData <- bind_rows(TheData_filtered)

#write.csv(TheData, file=StorageName, row.names=FALSE)
```

## QC_UserLibraries

When working with XML files for multiple users, the assembled dataframe output of `QC_LibraryParse()` can be passed to the `QC_UserLibraries` 
function, that will filter the individual library controls by user, and send them off to `Utility_Patchwork()` to create a .pdf output to the desired specifications. This can be useful tool for core staff to quickly generate outputs for each instrument user, that can then be sent to them to decide if they need to change any of their existing library reference controls. The saveCSV = TRUE argument would additionally return a .csv of the corresponding dataframe. 


```{r, eval=FALSE}
Outpath <- file.path("C:", "Users", "JohnDoe", "Desktop",
                             "LibraryControls5L")

TheIndividuals <- TheData %>% pull(Creator) %>% unique()
TheIndividuals

IndividualUser <- QC_UserLibraries(x=TheIndividuals[[1]], Data=TheData,
                                   NameAppend="_LibraryQC", outpath=Outpath,
                                   references = TRUE, thecolumns = 3, therows=4,
                                   width=7, height=9, saveCSV=FALSE)

AllUsers <- map(.x=TheIndividuals, .f=QC_UserLibraries, Data=TheData,
                 NameAppend="_LibraryQC", outpath=Outpath,references = TRUE,
                 thecolumns = 3, therows=4, width=7, height=9, saveCSV=FALSE)
```

# Comparing Normalized Signatures

We previously highlighted the ability within `QC_LibraryParse` to set reference=TRUE and compare our acquired library signatures to that of a reference signature. The following functions provide additionally functionality, extending this to signatures derrived from additional sources (including those derrived from .fcs files, covered extensively in Vignette 04_Fluorescent Signatures).

Let's for now continue by using a signature derrived from the Library controls highlighted above:

```{r}
LibraryData_Single <- QC_LibraryParse(XML_Files[2], returntype="dataframe", references=FALSE)
gt(LibraryData_Single)
```

As it currently exist, there are a few additional columns that are not required for the following functions, which we will remove. 
```{r}
Data <- LibraryData_Single %>% select(-Sample, -Creator, -Date) %>% rename(Sample=Fluorochrome)
gt(Data)
```

## QC_WhatsThis

The `QC_WhatsThis()` functions takes the above output, and will search the reference data for fluorophores with the closest resembling signature

```{r}
TheFluorophore <- Data %>% pull(Sample)

Results <- QC_WhatsThis(x=TheFluorophore, data=Data, NumberHits = 10, returnPlots=TRUE)

Results[[1]]
```

Setting returnPlots = TRUE, will additionally provide the plotted signatures. In combination with `ggplotly()`, this can be quite useful.

```{r}
plotly::ggplotly(Results[[2]])
```

As we can see, the acquired library signature differs from the reference signature. 

## QC_SimilarFluorophores

Additionally, we don't even need to provide a signature to compare to, if the fluorophore is present within the reference dataset, we can compare it to the other fluorophores present and similarly visualize it using the `QC_SimilarFluorophores()` option.

```{r}
Results <- QC_SimilarFluorophores(TheFluorophore="Spark Blue 550", NumberDetectors=64, NumberHits = 10, returnPlots=TRUE)

Results[[1]]
```

```{r}
plotly::ggplotly(Results[[2]])
```

## QC_ReferenceLibrary

If there is a mismatch in how the fluorophore is named within the reference data (or is absent), using the `QC_ReferenceLibrary()` function would be useful in correcting the inputed name for `QC_SimilarFluorophores()` for the output listed above. 


```{r}
QC_ReferenceLibrary(FluorNameContains = "Spark", NumberDetectors=64)
```

## QC_ProspectiveFluorophores

And finally, we have an experimental function `QC_ProspectiveFluorophores()` that will leverage information about an existing panel (and the peak detectors its fluorophore currently occupy) and contrast that to the reference library fluorophores to find candidate fluorophores that might potentially fit in the gaps. 

```{r}
OutPath <- file.path("C:", "Users", "JohnDoe", "Desktop")

Folder_Location <- system.file("extdata", package = "Luciernaga")
ThePanelLocation <- list.files(Folder_Location, pattern="^Panel.csv", full.names=TRUE)
ThePanel <- read.csv(ThePanelLocation, check.names=FALSE) %>% select(Fluorophore)
```

The arguments to provide to the function are a filepath to the panel.csv, NumberDetectors that your instrument has (64 for a 5L Aurora). TheCutoff corresponds to the cosine value when two fluorophores are compared to each other, I set the cutoff at 0.9 similarity. Rank Value comes from base R's `kappa()` how error introduced will spill over factoring across all fluorophore references in the matrix. Rough estimate for relative complexity in our case. 

```{r}
ProspectiveAdditions <- QC_ProspectiveAdditions(path=ThePanelLocation, NumberDetectors=64,
                                                TheCutoff=0.9, returnAll=FALSE, returnCSV=FALSE,
                                                filename="ProspectiveAdditions", outpath=OutPath)
gt(ProspectiveAdditions)
```

The results indicate potential fluorophores for Detectors not currently occupied by our panel, with cosine overlaps of less than TheCutoff (0.9). The RankValue is for the existing panel plus the given fluorophore. For panel design/exploratory purposes in our own lab, we would search for the Fluorophore name using `QC_SimilarFluorophores()` and identify other similar fluorophores by other commercial vendors to further investigate. 

For example from the above list:

```{r}
Results <- QC_SimilarFluorophores(TheFluorophore="PE-Fire 744", NumberDetectors=64, NumberHits = 10, returnPlots=TRUE)
plotly::ggplotly(Results[[2]])
```

# Concluding Thoughts

We hope that this vignette walkthrough was of interest, and that you can think of ways to implement it within your own workflow. If you have thought on how to improve things, please reach out! This was a particular fun section of the package to work on and I would love to continue to improve on the existing functions in the future. 


```{r sessionInfo, echo = FALSE}
sessionInfo()
```



